Multilevel generalizations of collective spin dynamics and squeezing

### 2020-04-28 collective_sun/multilevel_spin_notes.tex
changed some notation and added 'selection' rules for products of drive operators

\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\cov}{cov}
\let\var\relax
\DeclareMathOperator{\var}{var}

@@

Unlike the GGM operators in \eqref{eq:GGM_diag} and
\eqref{eq:GGM_off_diag}, the transition operators in
\eqref{eq:trans_ops} are not self-adjoint.  In order to express
Hamiltonians in terms of self-adjoint operators, we define the {\it
  drive operators}
\begin{align}
  D_{LM} \equiv \f{\eta_M}{\sqrt{2}}
  \sp{T_{L\abs{M}} + \sign\p{M} T_{L\abs{M}}^\dag},
  &&
  \eta_M \equiv
  \begin{cases}
    \sqrt{2} & M = 0 \\
    \p{-1}^M \sign\p{M}^{3/2} & M \ne 0
  \end{cases},
  \label{eq:drive_ops}
\end{align}
for integer $M$ with $\abs{M}\le L$.  The drive operator $D_{LM}$ is
proportional to the Hamiltonian induced on a nuclear spin by an
external classical field of spin-$L$ particles with spin projection
$M$ onto a quantization axis.  Similarly to the GGM operators in
\eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the drive operators
are proportional to the Pauli operators in the case of $n=2$, and
satisfy the orthonormality condition
\begin{align}
  \obk{D_{LM}|D_{L'M'}} = \delta_{LL'}\delta_{MM'},
\end{align}
which implies
\begin{align}
  H_{\t{int}} = -\f{u}{2} \sum_{p,q} \v D^{(p)}\c\v D^{(q)}
  = -\f{u}{2} \v\D \c \v\D,
  &&
  \v\D \equiv \sum_p \v D^{(p)},
  \label{eq:H_int_drive}
\end{align}
where $\v D$ is a vector of all drive operators in
\eqref{eq:drive_ops}.

@@

For our final re-expression of collective SU($n$)-symmetric
interactions, we abandon the use of single-spin operators entirely to
express the interaction Hamiltonian $H_{\t{int}}$ in terms of
operators that permute tensor factors (subsystems) of a Hilbert space
with an $N$-fold tensor product structure (e.g.~an array of nuclear
spins).  Specifically, we denote the Hilbert space of an $n$-level
system by $\H_n$, denote its $N$-fold tensor product by
$\H_n^{\otimes N}$, and identify the set $\Pi\equiv\set{\Pi_\sigma}$
of operators that permute tensor factors of $\H_n^{\otimes N}$
according to elements $\sigma\in\SS_N$ of the symmetric group of order
$N$.  We then expand
\begin{align}
  H_{\t{int}}
  = -\f{u}{2} \sum_{p,q,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  = -u\sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  - \f{u}{2} \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)},
  \label{eq:H_int_perm_start}
\end{align}
where the first sum,
\begin{align}
  \sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(q)}
  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu\nu}{\nu\mu}^{(p,q)}
  = \sum_{p<q} \Pi_{pq}
  \label{eq:spin_perm}
\end{align}
is simply a sum over all permutations $\Pi_{pq}$ of subsystems $p$ and
$q$; and the second sum,
\begin{align}
  \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)}
  = \sum_{p,\mu,\nu} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(p)}
  = n \sum_{p,\mu} \op{\mu}{\mu}^{(p)}
  = n \sum_p \1^{(p)}
  = n N,
  \label{eq:spin_const}
\end{align}
is merely a constant.  Substituting \eqref{eq:spin_perm} and
\eqref{eq:spin_const} into \eqref{eq:H_int_perm_start} yields
\begin{align}
  H_{\t{int}} = -u \sum_{p<q} \Pi_{pq} - \f12 N n u
  \simeq -u\sum_{p<q} \Pi_{pq},
  \label{eq:H_int_perm}
\end{align}
where $\simeq$ denotes equality up to an overall constant with no
physical consequence.  The form of the interaction Hamiltonian
$H_{\t{int}}$ makes it evident that its ground-state manifold
$\M_{\FS}$ is the set of {\it fully symmetric} (FS) states that have
eigenvalue 1 with respect to all pairwise permutation operators
$\Pi_{pq}$.  The fully symmetric manifold $\M_\FS$ is spanned by a
basis $\set{\ket{\v m}}$ of fully symmetric states that can be
uniquely identified by a list of integers
$\v m\equiv\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where $m_s$
indicates the total occupation number of subsystem state $s$, and
$\sum_sm_s=N$.  In the case of $n=2$, the fully symmetric manifold
$\M_\FS$ is precisely the Dicke manifold\cite{dicke1954coherence} of
$N$ qubits, spanned by Dicke states $\set{\ket{m_0,m_1}}$ for which
$m_0+m_1=N$ and the integer $m_0$ ($m_1$) indicates the total
occupation number of qubit state $\ket{0}$ ($\ket{1}$).  The dimension
of fully symmetric manifold $\M_\FS$ is determined by the number of
ways to assign $N$ subsystems to $n$ states, which is
${N+n-1 \choose n-1}$.  The energy of any fully symmetric state
$\ket{\psi_\FS}\in\M_\FS$ with respect to the interaction Hamiltonian
$H_{\t{int}}$ in \eqref{eq:H_int_perm} is
$-u{N \choose 2}=-\p{1/2}N\p{N-1}u$.

@@

Here we derive multiplication rules for transition operators on an $n$-dimensional Hilbert space, which allows us to expand
\begin{align}
  T_{\ell_1 m_1} T_{\ell_2 m_2}
  = \sum_{L,M} g_{\ell_1 m_1;\ell_2 m_2}^{LM} T_{LM},
\end{align}
with structure constants
\begin{align}
  g_{\ell_1 m_1;\ell_2 m_2}^{LM}
  \equiv \obk{T_{LM} | T_{\ell_1 m_1} T_{\ell_2 m_2}}
  = \tr\p{T_{LM}^\dag T_{\ell_1 m_1} T_{\ell_2 m_2}}.
  \label{eq:trans_struct}
\end{align}
For reference, the transition operators $T_{LM}$ are defined for
integers $L,M$ with $0\le L<n$ and $\abs{M}\le L$ in terms of
Clebsch-Gordan coefficients $\bk{\ell_1 m_1;\ell_2 m_2|\ell_3 m_3}$ by
\begin{align}
  T_{LM} \equiv \sqrt{\f{2L+1}{2I+1}}
  \sum_{\mu,\nu} \bk{I\mu;LM|I\nu} \op{\nu}{\mu},
  &&
  I \equiv \f{n-1}{2}.
\end{align}
Using the symmetry properties of Clebsch-Gordan coefficients, namely
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  &= \p{-1}^{\ell_2+m_2} \sqrt{\f{2\ell_3+1}{2\ell_1+1}}
  \bk{\ell_3,-m_3; \ell_2 m_2| \ell_1,-m_1} \\
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  &= \p{-1}^{\ell_1+\ell_2-\ell_3}
  \bk{\ell_1,-m_1; \ell_2,-m_2| \ell_3,-m_3},
\end{align}
we can find that
\begin{align}
  T_{LM}^\dag
  = \sqrt{\f{2L+1}{2I+1}}
  \sum_{\mu,\nu} \p{-1}^M \bk{I\nu;L,-M|I\mu} \op{\mu}{\nu}
  = \p{-1}^M T_{L,-M}.
\end{align}
Substituting this result into \eqref{eq:trans_struct} and expanding
Clebsch-Gordan coefficients in terms of Wigner 3-$j$ symbols as
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  = \p{-1}^{-\ell_1+\ell_2-m_3} \sqrt{\p{2\ell_3+1}}
  \begin{pmatrix}
    \ell_1 & \ell_2 & \ell_3 \\
    m_1 & m_2 & -m_3
  \end{pmatrix},
\end{align}
we find that
\begin{multline}
  g_{\ell_1 m_1;\ell_2 m_2}^{LM}
  = \delta_{M,m_1+m_2}
  \p{-1}^M \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}} \\
  \times \sum_{\mu,\nu,\rho} \p{-1}^{-3I+L+\ell_1+\ell_2-\mu-\nu-\rho}
  \begin{pmatrix}
    I & L & I \\
    \mu & -M & -\nu
  \end{pmatrix}
  \begin{pmatrix}
    I & \ell_1 & I \\
    \nu & m_1 & -\rho
  \end{pmatrix}
  \begin{pmatrix}
    I & \ell_2 & I \\
    \rho & m_2 & -\mu
  \end{pmatrix},
\end{multline}
where we can evaluate the sum and introduce Wigner 6-$j$ symbols to
get
\begin{align}
  g_{\ell_1 m_1;\ell_2 m_2}^{LM}
  &= \delta_{M,m_1+m_2} \p{-1}^{2I+M}
  \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}}
  \begin{pmatrix}
    L & \ell_1 & \ell_2 \\
    M & -m_1 & -m_2
  \end{pmatrix}
  \begin{Bmatrix}
    L & \ell_1 & \ell_2 \\
    I & I & I
  \end{Bmatrix} \\
  &= \delta_{M,m_1+m_2} \p{-1}^{2I+L}
  \sqrt{\p{2\ell_1+1}\p{2\ell_2+1}}
  \bk{\ell_1 m_1; \ell_2 m_2| LM}
  \begin{Bmatrix}
    \ell_1 & \ell_2 & L \\
    I & I & I
  \end{Bmatrix}.
  \label{eq:trans_prod}
\end{align}
The Wigner 6-$j$ symbol is symmetric under any permutation of its
columns, while Clebsch-Gordan coefficients satisfy
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  = \p{-1}^{\ell_1+\ell_2-\ell_3}
  \bk{\ell_2 m_2; \ell_1 m_1| \ell_3 m_3},
\end{align}
which implies that
\begin{align}
  g_{\ell_1 m_1;\ell_2 m_2}^{LM}
  = \p{-1}^{\ell_1+\ell_2-L} g_{\ell_2 m_2;\ell_1 m_1}^{LM}.
\end{align}

@@

The drive operators are defined by
\begin{align}
  D_{LM} \equiv \f{\eta_M}{\sqrt{2}}
  \sp{T_{L\abs{M}} + \sign\p{M} T_{L\abs{M}}^\dag},
  &&
  \eta_M \equiv
  \begin{cases}
    \sqrt{2} & M = 0 \\
    \p{-1}^M \sign\p{M}^{3/2} & M \ne 0
  \end{cases},
\end{align}
or, equivalently,
\begin{align}
  D_{LM} \equiv \f{\tau_M}{\sqrt{2}}
  \sp{T_{LM} + \sign\p{M} T_{LM}^\dag},
  &&
  \tau_M \equiv
  \begin{cases}
    \sqrt{2} & M = 0 \\
    \p{-1}^M \sign\p{M}^{M+1/2} & M \ne 0
  \end{cases},
\end{align}
which implies that
\begin{align}
  T_{LM} = \f{\tilde\tau_M}{\sqrt{2}} \p{D_{LM} + i D_{L,-M}},
  &&
  \tilde\tau_M \equiv
  \begin{cases}
    \sqrt{2}/\p{1+i} & M = 0 \\
    \tau_M^* & M \ne 0
  \end{cases},
\end{align}
where $\tau_M^*$ is the complex conjugate of $\tau_M$.  In order to
evaluate the product of two drive operators, we can thus expand them
in terms of transition operators, evaluate the corresponding products
of transition operators (see Appendix \ref{sec:trans_prod}), and
finally expand the result in terms of drive operators.  The general
product of two drive operators takes the form
\begin{align}
  D_{\ell_1 m_1} D_{\ell_2 m_2}
  = \f{\tau_{m_1}\tau_{m_2}}{\sqrt{2}}
  \sum_L \p{g_{\ell_1 m_1;\ell_2 m_2}^{L,m_1+m_2}
    E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
    + g_{\ell_1 m_1;\ell_2,-m_2}^{L,m_1-m_2}
    E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}},
\end{align}
where in terms of
\begin{align}
  \epsilon_M \equiv \p{-1}^M \sign\p{M},
\end{align}
we have
\begin{align}
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
  &\equiv \sum_{s\in\set{\pm1}}
  \f12 \sp{\tilde\tau_{m_1+m_2}
    + i s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \epsilon_{m_2} \tilde\tau_{-\p{m_1+m_2}}}
  \sqrt{s} D_{L,s\p{m_1+m_2}},
  \\
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}
  &\equiv \sum_{s\in\set{\pm1}}
  \f12 \sp{\epsilon_{m_2} \tilde\tau_{m_1-m_2}
    + i s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \tilde\tau_{-\p{m_1-m_2}}}
  \sqrt{s} D_{L,s\p{m_1-m_2}}.
\end{align}
If $m_1\pm m_2=0$, these operators simplify to
\begin{align}
  E_{\ell_1 m;\ell_2,-m}^{(L,+)}
  &= \f1{\sqrt{2}} \sp{1 - \p{-1}^{L+\ell_1+\ell_2}
    \p{1-\delta_{m,0}}} D_{L,0}, \\
  E_{\ell_1 m;\ell_2 m}^{(L,-)}
  &= \f{\epsilon_m}{\sqrt{2}}
  \sp{1 + \p{-1}^{L+\ell_1+\ell_2}} D_{L,0},
\end{align}
and are otherwise
\begin{align}
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
  &\stackrel{m_1+m_2\ne0}{=}
  \tau_{m_1+m_2}^* \sum_{s\in\set{\pm1}}
  \f12 \sp{1 + s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \epsilon_{m_2} \epsilon_{m_1+m_2}}
  \sqrt{s} D_{L,s\p{m_1+m_2}}, \\
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}
  &\stackrel{m_1-m_2\ne0}{=}
  \tau_{m_1-m_2}^* \sum_{s\in\set{\pm1}}
  \f12 \sp{\epsilon_{m_2} + s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \epsilon_{m_1-m_2}}
  \sqrt{s} D_{L,s\p{m_1-m_2}}.
\end{align}
If $\p{m_1,m_2}=\p{0,m}$, then we can use the fact that
$i\tau_m\tau_{-m}^*=\epsilon_m$ for $m\ne0$ to simplify
\begin{align}
  D_{\ell_1,0} D_{\ell_2 m}
  = \sum_L g_{\ell_1,0,\ell_2,m}^{Lm}
  \sum_{s\in\set{\pm1}} \f12 \sp{1 + s \p{-1}^{L+\ell_1+\ell_2}}
  \sqrt{s} D_{L,sm}.
\end{align}

>>>>

\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\cov}{cov}
\let\var\relax
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\even}{even}

@@

Unlike the GGM operators in \eqref{eq:GGM_diag} and
\eqref{eq:GGM_off_diag}, the transition operators in
\eqref{eq:trans_ops} are not self-adjoint.  In order to express
Hamiltonians in terms of self-adjoint operators, we define the {\it
  drive operators}
\begin{align}
  D_{LM} \equiv \f{\eta_M}{\sqrt{2}}
  \sp{T_{LM} + \sign\p{M} T_{LM}^\dag},
  &&
  \eta_M \equiv
  \begin{cases}
    \sqrt{2} & M = 0 \\
    \p{-1}^M \sign\p{M}^{M+1/2} & M \ne 0
  \end{cases},
  \label{eq:drive_ops}
\end{align}
for integer $M$ with $\abs{M}\le L$.  The drive operator $D_{LM}$ is
proportional to the Hamiltonian induced on a nuclear spin by an
external classical field of spin-$L$ particles with spin projection
$M$ onto a quantization axis.  Similarly to the GGM operators in
\eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the drive operators
are proportional to the Pauli operators in the case of $n=2$, and
satisfy the orthonormality condition
\begin{align}
  \obk{D_{LM}|D_{L'M'}} = \delta_{LL'}\delta_{MM'},
\end{align}
which implies
\begin{align}
  H_{\t{int}} = -\f{u}{2} \sum_{p,q} \v D^{(p)}\c\v D^{(q)}
  = -\f{u}{2} \v\D \c \v\D,
  &&
  \v\D \equiv \sum_p \v D^{(p)},
  \label{eq:H_int_drive}
\end{align}
where $\v D$ is a vector of all drive operators in
\eqref{eq:drive_ops}.

@@

For our final re-expression of collective SU($n$)-symmetric
interactions, we abandon the use of single-spin operators entirely to
express the interaction Hamiltonian $H_{\t{int}}$ in terms of
operators that permute tensor factors (subsystems) of a Hilbert space
with an $N$-fold tensor product structure (e.g.~an array of nuclear
spins).  Specifically, we denote the Hilbert space of an $n$-level
system by $\H_n$, denote its $N$-fold tensor product by
$\H_n^{\otimes N}$, and identify the set $\Pi\equiv\set{\Pi_s}$ of
operators that permute tensor factors of $\H_n^{\otimes N}$ according
to elements $s\in\SS_N$ of the symmetric group of order $N$.  We then
expand
\begin{align}
  H_{\t{int}}
  = -\f{u}{2} \sum_{p,q,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  = -u\sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  - \f{u}{2} \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)},
  \label{eq:H_int_perm_start}
\end{align}
where the first sum,
\begin{align}
  \sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(q)}
  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu\nu}{\nu\mu}^{(p,q)}
  = \sum_{p<q} \Pi_{pq}
  \label{eq:spin_perm}
\end{align}
is simply a sum over all permutations $\Pi_{pq}$ of subsystems $p$ and
$q$; and the second sum,
\begin{align}
  \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)}
  = \sum_{p,\mu,\nu} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(p)}
  = n \sum_{p,\mu} \op{\mu}{\mu}^{(p)}
  = n \sum_p \1^{(p)}
  = n N,
  \label{eq:spin_const}
\end{align}
is merely a constant.  Substituting \eqref{eq:spin_perm} and
\eqref{eq:spin_const} into \eqref{eq:H_int_perm_start} yields
\begin{align}
  H_{\t{int}} = -u \sum_{p<q} \Pi_{pq} - \f12 N n u
  \simeq -u\sum_{p<q} \Pi_{pq},
  \label{eq:H_int_perm}
\end{align}
where $\simeq$ denotes equality up to an overall constant with no
physical consequence.  The form of the interaction Hamiltonian
$H_{\t{int}}$ makes it evident that its ground-state manifold
$\M_{\FS}$ is the set of {\it fully symmetric} (FS) states that have
eigenvalue 1 with respect to all pairwise permutation operators
$\Pi_{pq}$.  The fully symmetric manifold $\M_\FS$ is spanned by a
basis $\set{\ket{\v m}}$ of fully symmetric states that can be
uniquely identified by a list of integers
$\v m\equiv\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where $m_s$
indicates the total occupation number of subsystem state $s$, and
$\sum_sm_s=N$.  In the case of $n=2$, the fully symmetric manifold
$\M_\FS$ is precisely the Dicke manifold\cite{dicke1954coherence} of
$N$ qubits, spanned by Dicke states $\set{\ket{m_0,m_1}}$ for which
$m_0+m_1=N$ and the integer $m_0$ ($m_1$) indicates the total
occupation number of qubit state $\ket{0}$ ($\ket{1}$).  The dimension
of fully symmetric manifold $\M_\FS$ is determined by the number of
ways to assign $N$ subsystems to $n$ states, which is
${N+n-1 \choose n-1}$.  The energy of any fully symmetric state
$\ket{\psi_\FS}\in\M_\FS$ with respect to the interaction Hamiltonian
$H_{\t{int}}$ in \eqref{eq:H_int_perm} is
$-u{N \choose 2}=-\p{1/2}N\p{N-1}u$.

@@

Here we derive multiplication rules for transition operators on an $n$-dimensional Hilbert space, which allows us to expand
\begin{align}
  T_{\ell_1 m_1} T_{\ell_2 m_2}
  = \sum_{L,M} f_{\ell_1 m_1;\ell_2 m_2}^{LM} T_{LM},
\end{align}
with structure constants
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  \equiv \obk{T_{LM} | T_{\ell_1 m_1} T_{\ell_2 m_2}}
  = \tr\p{T_{LM}^\dag T_{\ell_1 m_1} T_{\ell_2 m_2}}.
  \label{eq:trans_struct_start}
\end{align}
For reference, the transition operators $T_{LM}$ are defined for
integers $L,M$ with $0\le L<n$ and $\abs{M}\le L$ in terms of
Clebsch-Gordan coefficients $\bk{\ell_1 m_1;\ell_2 m_2|\ell_3 m_3}$ by
\begin{align}
  T_{LM} \equiv \sqrt{\f{2L+1}{2I+1}}
  \sum_{\mu,\nu} \bk{I\mu;LM|I\nu} \op{\nu}{\mu},
  &&
  I \equiv \f{n-1}{2}.
\end{align}
Using the symmetry properties of Clebsch-Gordan coefficients, namely
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  &= \p{-1}^{\ell_2+m_2} \sqrt{\f{2\ell_3+1}{2\ell_1+1}}
  \bk{\ell_3,-m_3; \ell_2 m_2| \ell_1,-m_1} \\
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  &= \p{-1}^{\ell_1+\ell_2-\ell_3}
  \bk{\ell_1,-m_1; \ell_2,-m_2| \ell_3,-m_3},
\end{align}
we can find that
\begin{align}
  T_{LM}^\dag
  = \sqrt{\f{2L+1}{2I+1}}
  \sum_{\mu,\nu} \p{-1}^M \bk{I\nu;L,-M|I\mu} \op{\mu}{\nu}
  = \p{-1}^M T_{L,-M}.
\end{align}
Substituting this result into \eqref{eq:trans_struct_start} and
expanding Clebsch-Gordan coefficients in terms of Wigner 3-$j$ symbols
as
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  = \p{-1}^{-\ell_1+\ell_2-m_3} \sqrt{\p{2\ell_3+1}}
  \begin{pmatrix}
    \ell_1 & \ell_2 & \ell_3 \\
    m_1 & m_2 & -m_3
  \end{pmatrix},
\end{align}
we find that
\begin{multline}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  = \delta_{M,m_1+m_2}
  \p{-1}^M \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}} \\
  \times \sum_{\mu,\nu,\rho} \p{-1}^{-3I+L+\ell_1+\ell_2-\mu-\nu-\rho}
  \begin{pmatrix}
    I & L & I \\
    \mu & -M & -\nu
  \end{pmatrix}
  \begin{pmatrix}
    I & \ell_1 & I \\
    \nu & m_1 & -\rho
  \end{pmatrix}
  \begin{pmatrix}
    I & \ell_2 & I \\
    \rho & m_2 & -\mu
  \end{pmatrix},
\end{multline}
where we can evaluate the sum and introduce Wigner 6-$j$ symbols to
get
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  &= \delta_{M,m_1+m_2} \p{-1}^{2I+M}
  \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}}
  \begin{pmatrix}
    L & \ell_1 & \ell_2 \\
    M & -m_1 & -m_2
  \end{pmatrix}
  \begin{Bmatrix}
    L & \ell_1 & \ell_2 \\
    I & I & I
  \end{Bmatrix} \\
  &= \delta_{M,m_1+m_2} \p{-1}^{2I+L}
  \sqrt{\p{2\ell_1+1}\p{2\ell_2+1}}
  \bk{\ell_1 m_1; \ell_2 m_2| LM}
  \begin{Bmatrix}
    \ell_1 & \ell_2 & L \\
    I & I & I
  \end{Bmatrix}.
  \label{eq:trans_struct}
\end{align}
The Wigner 6-$j$ symbol is symmetric under any permutation of its
columns, while Clebsch-Gordan coefficients satisfy
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  &= \p{-1}^{\ell_1+\ell_2-\ell_3}
  \bk{\ell_1,-m_1; \ell_2, -m_2| \ell_3 m_3} \\
  &= \p{-1}^{\ell_1+\ell_2-\ell_3}
  \bk{\ell_2 m_2; \ell_1 m_1| \ell_3 m_3},
\end{align}
which implies that
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  &= \p{-1}^{\ell_1+\ell_2-L} f_{\ell_1,-m_2;\ell_2,-m_2}^{L,-M} \\
  &= \p{-1}^{\ell_1+\ell_2-L} f_{\ell_2 m_2;\ell_1 m_1}^{LM}.
\end{align}

@@

The drive operators are defined by
\begin{align}
  D_{LM} \equiv \f{\eta_M}{\sqrt{2}}
  \sp{T_{LM} + \sign\p{M} T_{LM}^\dag},
  &&
  \eta_M \equiv
  \begin{cases}
    \sqrt{2} & M = 0 \\
    \p{-1}^M \sign\p{M}^{M+1/2} & M \ne 0
  \end{cases},
\end{align}
or, equivalently,
\begin{align}
  D_{LM} \equiv \f{\eta_M}{\sqrt{2}}
  \sum_{s\in\set{\pm1}} \epsilon_{Ms} T_{L,sM},
  &&
  \epsilon_{Ms} \equiv
  \begin{cases}
    1 & s = +1 \\
    \p{-1}^M \sign\p{M} & s = -1
  \end{cases},
\end{align}
which implies that
\begin{align}
  T_{LM} = \f{\tau_M}{\sqrt{2}}
  \sum_{s\in\set{\pm1}} \sqrt{s} D_{L,sM},
  &&
  \tau_M \equiv
  \begin{cases}
    \sqrt{2}/\p{1+i} & M = 0 \\
    \eta_M^* & M \ne 0
  \end{cases},
\end{align}
where $\eta_M^*$ is the complex conjugate of $\eta_M$.  In order to
evaluate the product of two drive operators, we can thus expand them
in terms of transition operators, evaluate the corresponding products
of transition operators (see Appendix \ref{sec:trans_prod}), and
finally expand the result in terms of drive operators.  The product of
two drive operators is thus
\begin{align}
  D_{\ell_1 m_1} D_{\ell_2 m_2}
  = \sum_{L,M} g_{\ell_1m_1;\ell_2m_2}^{LM} D_{LM}
  \label{eq:drive_prod}
\end{align}
where
\begin{align}
  g_{\ell_1m_1;\ell_2m_2}^{LM}
  &\equiv \f{\eta_{m_1}\eta_{m_2}}{2\sqrt2}
  \sum_{r,s_1,s_2\in\set{\pm1}} \epsilon_{m_1s_1} \epsilon_{m_2s_2}
  \times f_{\ell_1,s_1m_1;\ell_2,s_2m_2}^{L,rM}
  \times \tau_{s_1m_1+s_2m_2} \times \sqrt{r}.
\end{align}
Redefining $\p{s_1,s_2}\to\p{rs_1,rs_2}$ and using the facts that
\begin{align}
  f_{\ell_1,rs_1m_1;\ell_2,rs_2m_2}^{L,rM}
  = r^{\ell_1+\ell_2+L} f_{\ell_1,s_1m_1;\ell_2,s_2m_2}^{LM},
  &&
  \tau_{rM} \sqrt{r} &= \tau_M \, \sigma_{Mr} \, r^M,
\end{align}
with
\begin{align}
  \sigma_{Mr} \equiv
  \begin{cases}
    \sqrt{r} & M = 0 \\
    \sign\p{M}^{\frac{1-r}{2}} & M \ne 0
  \end{cases},
  &&
  0^0 \equiv 1,
\end{align}
we can simplify
\begin{align}
  g_{\ell_1m_1;\ell_2m_2}^{LM}
  = \f{\eta_{m_1}\eta_{m_2}\tau_M}{\sqrt2}
  \sum_{s_1,s_2\in\set{\pm1}} f_{\ell_1,s_1m_1;\ell_2,s_2m_2}^{LM}
  \times \f12 \sum_{r\in\set{\pm1}}
  \epsilon_{m_1,r s_1} \epsilon_{m_2,r s_2} \,
  \sigma_{Mr} \, r^{\ell_1+\ell_2+L+M}.
  \label{eq:drive_struct}
\end{align}
In the case that $m_1=0$, we can simplify
\begin{align}
  g_{\ell_1,0;\ell_2,m}^{LM}
  \stackrel{m\ne 0}{=} f_{\ell_1,0;\ell_2 m}^{Lm}
  \times \delta_{\abs{M},\abs{m}} \times
  \begin{cases}
    \sp{\sp{\ell_1+\ell_2+L~\t{is even}}} & M = +m, \\
    i \sp{\sp{\ell_1+\ell_2+L~\t{is odd}}} & M = -m,
  \end{cases}
\end{align}
\begin{align}
  g_{\ell_1,0;\ell_2,0}^{LM} = f_{\ell_1,0;\ell_2,0}^{LM}.
\end{align}
Here $\sp{\sp{X}}\equiv1$ if $X$ is true and $0$ otherwise.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Finding drive operator sub-algebras}
